---
title: "SSURGO all counties"
author: "Luke Salvato"
date: "1/19/2021"
output: html_document
---


```{r setup, include=FALSE}
library(FedData)
library(sf)
library(tidyverse)

#create and save function
not_all_na <- function(x) any(!is.na(x)) #for data cleaning

getwd()
```

```{r}
#Colusa County
SSURGO.areas.011 <- get_ssurgo(template=c('CA011'), 
                               label='CA_011')

#Glen County
SSURGO.areas.021 <- get_ssurgo(template=c('CA021'), 
                               label='CA_021')

#Solano County
SSURGO.areas.095 <- get_ssurgo(template=c('CA095'), 
                               label='CA_095')

#Yolo County
SSURGO.areas.113 <- get_ssurgo(template=c('CA113'), 
                               label='CA_113')

#Butte County
SSURGO.areas.612 <- get_ssurgo(template=c('CA612'), 
                               label='CA_612')
getwd()


#Sutter County
SSURGO.areas.101 <- get_ssurgo(template=c('CA101'), 
                              label='CA_101') #This one not working for some reason

#Sacramento County
SSURGO.areas.067 <- get_ssurgo(template=c('CA067'), 
                              label='CA_067') #This one also is not working for some reason

#Placer County
SSURGO.areas.620 <- get_ssurgo(template=c('CA620'), 
                               label='CA_620')

#Yuba County
SSURGO.areas.618 <- get_ssurgo(template=c('CA618'), 
                               label='CA_618')

#Tehema County
SSURGO.areas.645 <- get_ssurgo(template=c('CA645'), 
                               label='CA_645')
```

```{r}

chorizon <- rbind(SSURGO.areas.011$tabular$chorizon, SSURGO.areas.021$tabular$chorizon, SSURGO.areas.095$tabular$chorizon, SSURGO.areas.113$tabular$chorizon, SSURGO.areas.612$tabular$chorizon, SSURGO.areas.101$tabular$chorizon, SSURGO.areas.067$tabular$chorizon, SSURGO.areas.620$tabular$chorizon, SSURGO.areas.618$tabular$chorizon, SSURGO.areas.645$tabular$chorizon)


component <- rbind(SSURGO.areas.011$tabular$component, SSURGO.areas.021$tabular$component, SSURGO.areas.095$tabular$component, SSURGO.areas.113$tabular$component, SSURGO.areas.612$tabular$component, SSURGO.areas.101$tabular$component, SSURGO.areas.067$tabular$component, SSURGO.areas.620$tabular$component, SSURGO.areas.618$tabular$component, SSURGO.areas.645$tabular$component)


mapunit <- rbind(SSURGO.areas.011$tabular$mapunit, SSURGO.areas.021$tabular$mapunit, SSURGO.areas.095$tabular$mapunit, SSURGO.areas.113$tabular$mapunit, SSURGO.areas.612$tabular$mapunit, SSURGO.areas.101$tabular$mapunit, SSURGO.areas.067$tabular$mapunit, SSURGO.areas.620$tabular$mapunit, SSURGO.areas.618$tabular$mapunit, SSURGO.areas.645$tabular$mapunit)

corestrictions <- rbind(SSURGO.areas.011$tabular$corestrictions, SSURGO.areas.021$tabular$corestrictions, SSURGO.areas.095$tabular$corestrictions, SSURGO.areas.113$tabular$corestrictions, SSURGO.areas.612$tabular$corestrictions, SSURGO.areas.101$tabular$corestrictions, SSURGO.areas.067$tabular$corestrictions, SSURGO.areas.620$tabular$corestrictions, SSURGO.areas.618$tabular$corestrictions, SSURGO.areas.645$tabular$corestrictions)

corestrictions %>% count(reskind)
corestrictions %>% count(reshard)





```


Remove columns that have only NAs (using the function I built above)
```{r}
chorizon <- chorizon %>% 
  select_if(not_all_na)
component <- component %>% 
  select_if(not_all_na)
mapunit <- mapunit %>%
  select_if(not_all_na)
corestrictions <- corestrictions %>%
  select_if(not_all_na)
```

Create a new dataframe that consists of the deepest horizon bottom for each cokey
```{r}
#deepest horizon bottom of each component
depth <- chorizon %>%
  group_by(cokey) %>%
  summarise(total.depth = max(hzdepb.r))
head(depth)

#remove horizons that start below 30 cm
chorizon <- chorizon %>%
  filter(hzdept.r < 30) %>% 
  droplevels()
```

Wil combine chorizon to the component based on cokey 
We want only one observation per cokey for the join from chorizon data to component data
(we will then combine component_horizon data to the mapunit data based on mukey)
```{r}
component_count <- chorizon %>%
  group_by(cokey) %>%
  summarize(count = n()) %>%
  filter(count > 1)

nrow(component_count)
#there are 1972 compnents with more than one horizon layer in the first 30cm
```

Summarize characteristics of interest with a weighted mean of the horizons present
```{r}
#create a variable thick which will be used to relate the amount of soil in each horizon in each component
chorizon_vars <- chorizon %>%
  mutate(thick = ifelse(hzdepb.r > 30, 30 - hzdept.r, 
                        hzdepb.r - hzdept.r)) %>% 
  group_by(cokey) %>%
  summarise(sand = round(weighted.mean(sandtotal.r, thick, na.rm = TRUE),2),
            silt = round(weighted.mean(silttotal.r, thick, na.rm = TRUE),2),
            clay = round(weighted.mean(claytotal.r, thick, na.rm = TRUE),2),
            om = round(weighted.mean(om.r, thick, na.rm = TRUE),2),
            k = round(weighted.mean(kffact, thick, na.rm = TRUE),2),
            cec = round(weighted.mean(cec7.r, thick, na.rm = TRUE),2),
            ph = round(weighted.mean(ph1to1h2o.r, thick),2),
            ec = round(weighted.mean(ec.r, thick, na.rm = TRUE),2),
            lep = round(weighted.mean(lep.r, thick, na.rm = TRUE),2),
            sar = round(weighted.mean(sar.r, thick, na.rm = TRUE),2),
            ksat.r_mean = round(weighted.mean(ksat.r, thick, na.rm = TRUE),2),
            ksat.r_min = round(min(ksat.r, thick, na.rm = TRUE),2),
            ksat.r_max = round(max(ksat.r, thick, na.rm = TRUE),2),
            ksat.l_mean = round(weighted.mean(ksat.l, thick, na.rm = TRUE),2),
            ksat.l_min = round(min(ksat.l, thick, na.rm = TRUE),2),
            ksat.h_mean = round(weighted.mean(ksat.h, thick, na.rm = TRUE),2),
            ksat.h_max = round(max(ksat.h, thick, na.rm = TRUE),2))
            
#now we have one observation per component for each of the important variables
```

```{r}
chorizon_vars_depth <- left_join(chorizon_vars, depth, by = "cokey")
head(chorizon)
#Note: the row with all NAs is possibly data for a rock outcrop mapunit? can vie this later when we attach chorizon and component
```

```{r}
#Inspect the column names of df 'component'
length(colnames(component))
colnames(component)
```
component data: column selection
```{r}
#use dplyr select to select the important columns
component <- component %>%
  dplyr::select(c(comppct.r, compname, majcompflag, slope.r, 
                  slopelenusle.r, runoff, tfact, wei, weg, erocl, 
                  elev.r, albedodry.r, airtempa.r, map.r, ffd.r, compname,
                taxorder, irrcapcl,
                  #cropprodindex, #for some reason this one doesn't exist in this data set
                  taxpartsize, mukey, cokey, drainagecl)) %>% 
  mutate(MUKEY = as.character(mukey))
dim(component)
```

```{r}
component_horizon <- left_join(component, chorizon_vars_depth, by = c("cokey")) #%>% 
  

component_horizon_corestrictions <- left_join(component_horizon,corestrictions, by = c("cokey"))
dim(component_horizon_corestrictions)

#joined and cleaned:
head(component_horizon_corestrictions)
```

Mapunit data: column selection
```{r}
colnames(mapunit)

mapunit <- mapunit %>%
  dplyr::select(c(musym, muname, muacres, mukey, farmlndcl))#, county))

```
Join component_horizon with mapunit
This time join using 'mukey' - get component level information for each mapunit
recall that we will have some mapunits without data because there was no horizon level data
```{r}
full_soil <- left_join(component_horizon_corestrictions, mapunit, by = c("mukey"))
full_soil <- full_soil %>%
  mutate(muname = gsub(", ", "_", muname)) %>% 
  mutate(MUKEY = as.character(mukey)) 
  
```




Make area weighted average for each MUKEY
```{r}
awav <- full_soil %>% 
  group_by(MUKEY) %>% 
  summarise(sand = round(weighted.mean(sand, comppct.r, na.rm = TRUE),2),
            silt = round(weighted.mean(silt, comppct.r, na.rm = TRUE),2),
            clay = round(weighted.mean(clay, comppct.r, na.rm = TRUE),2),
            om = round(weighted.mean(om, comppct.r, na.rm = TRUE),2),
            k = round(weighted.mean(k, comppct.r, na.rm = TRUE),2),
            cec = round(weighted.mean(cec, comppct.r, na.rm = TRUE),2),
            ph = round(weighted.mean(ph, comppct.r, na.rm = TRUE),2),
            lep = round(weighted.mean(lep, comppct.r, na.rm = TRUE),2),
            sar = round(weighted.mean(sar, comppct.r, na.rm = TRUE),2),
            ksat.r_mean = round(weighted.mean(ksat.r_mean, comppct.r, na.rm = TRUE),2),
            ksat.r_min = round(weighted.mean(ksat.r_min, comppct.r, na.rm = TRUE),2),
            ksat.r_max = round(weighted.mean(ksat.r_max, comppct.r, na.rm = TRUE),2),
            ksat.l_mean = round(weighted.mean(ksat.l_mean, comppct.r, na.rm = TRUE),2),
            ksat.l_min = round(weighted.mean(ksat.l_min, comppct.r, na.rm = TRUE),2),
            ksat.h_mean = round(weighted.mean(ksat.h_mean, comppct.r, na.rm = TRUE),2),
            ksat.h_max = round(weighted.mean(ksat.h_max, comppct.r, na.rm = TRUE),2),) %>% 
  ungroup()



names(awav)
  
``` 
`
```{r}
lean_soil <- component_horizon_corestrictions %>% 
  select(-c(sand, silt,clay,om,k,cec,ph,lep,sar,ksat.r_mean, ksat.r_min, ksat.r_max, ksat.l_mean, ksat.l_min, ksat.h_mean, ksat.h_max )) %>% 
  right_join(mapunit, by = c("mukey")) %>% 
  mutate(muname = gsub(", ", "_", muname)) %>% 
  mutate(MUKEY = as.character(mukey)) %>% 
  right_join(awav, by = "MUKEY") 
  #left_join()
  

```

Make a 'yes/no' label for restrictive layers
Also tak eonly the most dominant component, using comppct.r (precent component )
```{r}
lean_soil_dominant_componant <- lean_soil %>% 
  mutate(restrictive_layer = case_when(reshard == c("Indurated", "Strongly cemented", "Very strongly cemented") ~ "yes",
                                       TRUE ~ "no")) %>% 
  filter(majcompflag == "Yes") %>% 
  group_by(mukey) %>% 
  slice(which.max(comppct.r)) 

  
#lean_soil_dominant_componant %>% count(MUKEY)  
#should be one of each mukey now
```


Now need to incorporate the spatial data, so that I can make maps
Join full_soil with spatial data
```{r}
sf <- rbind(SSURGO.areas.011$spatial, SSURGO.areas.021$spatial, SSURGO.areas.095$spatial, SSURGO.areas.113$spatial, SSURGO.areas.612$spatial, SSURGO.areas.101$spatial, SSURGO.areas.067$spatial, SSURGO.areas.620$spatial, SSURGO.areas.618$spatial, SSURGO.areas.645$spatial) %>% 
  st_as_sf() %>% 
  mutate(MUKEY = as.character(MUKEY))%>% 
 left_join(lean_soil_dominant_componant, by = "MUKEY") %>% 
  filter(majcompflag == "Yes") %>% 
  dplyr::select(MUKEY, comppct.r, compname, taxorder, farmlndcl, irrcapcl, cokey, drainagecl, total.depth, reskind,reshard, ec, sand, silt, clay, om, cec, ph, lep, sar, muacres, muname, restrictive_layer, ksat.r_mean, ksat.r_min, ksat.r_max, ksat.l_mean, ksat.l_min, ksat.h_mean, ksat.h_max )

sf %>% as_tibble() %>% count(MUKEY)
```

```{r} 
#Plot counties clay content
ggplot() + geom_sf(data = sf, aes(fill = clay), size = 0.00000005)

```
Inspect NAs
```{r}
#Convert the sf object back to a tibble to inspect these NAs
soil_tb <- sf %>% as_tibble()
#count the NAs in each column, try to see where the NAs are
where_NA <- map(soil_tb, ~sum(is.na(.)))
glimpse(where_NA)



#where the NAs are coming from? .
library(janitor)
where_na <- sf %>% 
  as_tibble %>%
  filter(is.na(clay)) %>% 
  select(sand, silt, clay, compname) %>%
  tabyl(clay, compname)
where_na

```

Write the shape file
```{r}
getwd()
#st_write(sf, "all_counties_ssurgo_may_18_2021.shp", delete_layer = TRUE)
#st_write(sf, "all_counties_ssurgo_20210922.geojson")
st_write(sf, "/Users/lukesalvato/Dropbox/Mac (3)/Documents/Land_Use/soils/data/all_counties_ssurgo_20220209.geojson", delete_layer = TRUE)
```


Plot to visualize ksat
```{r}
soil_tb %>% ggplot(aes(x = ksat.r_mean))+
  geom_density()
soil_tb %>% ggplot(aes(x = ksat.l_min))+
  geom_density()

soil_tb %>% ggplot(aes(x = lep))+
  geom_density()
soil_tb %>% ggplot(aes(x = sar))+
  geom_density()
```

